% SVN info for this file
%!TEX root = report.tex

\svnidlong
{$HeadURL$}
{$LastChangedDate$}
{$LastChangedRevision$}
{$LastChangedBy$}

\chapter{Valorization of Work}
\labelChapter{value}

\lettrine{D}{istribution} System Operators (DSOs) are facing more problems by the day with the increasing penetration of DRES. So far, research has not be able to sufficiently identify solutions for this problem, or even assuage it. The little amount of research that attempts to treat this problem has several shortcomings that will render it unusable in real-world networks.\\

The research work proposed here has the provides the following value additions when compared to the state-of-the-art in the field:

% \begin{itemize}
% \item This work has the potential to be a part of one of the first solutions to completely treat the problems, as outlined in the report, faced by DSOs today, and is a part of a pan-European project that has clear, marketable, and relevant objectives.\\

% \item By using economic models to provide a solution for DSOs with the lowest cost, all the constraints have been brought on to a single scale of reference, cost. This makes it very interesting from a business point of view and not just from an academic research point of view.\\

% \item Unlike most research, this work, including what has been developed and will be developed in the future, has remained and will strive to remain as close to real-world conditions as possible. This will permit an easy transition for the resulting methods from the drawing board to their potential applications in the real-world.\\

% \end{itemize}

\section*{A Complete Solution}
This work has the potential to be a part of one of the first solutions to completely treat the problems, as outlined in the report, faced by DSOs today, and is a part of a pan-European project that has clear, marketable, and relevant objectives.\\

With all the envisaged developments, once the work is complete, it will result in a cutting-edge application that can be directly applied to any distribution network for day-ahead optimization. And with the solution being modular, changes and upgrades can be brought about with ease.

\section*{Research that is Applicable}
Unlike most research, this work, including what has been developed and will be developed in the future, has remained and will strive to remain as close to real-world conditions as possible. This will permit an easy transition for the resulting methods from the drawing board to their potential applications in the real-world.

\section*{The Most Economical Solution}
Most current research focuses on objectives such as loss reduction, or consumer satisfaction, while not being able to balance this with other objectives.\\

By using economic models to provide a solution with the lowest cost for DSOs, all the constraints and objectives have been brought on to a single scale of reference: cost. This makes it very interesting from a business point of view and not just from an academic research point of view. The results obtained from distribution networks will, provided detailed economic models are constructed, be the most interesting solution to DSOs. Even with assumptions for these models, the results are encouraging.
